Visualizing Multivariate PDFs

### 2016-08-24 sections/seperable.tex
functional box

Because the box-plot does not give any indication of how data is distributed
inside the box or along the whiskers, the histplot and vaseplot were introduced to indicate
the density of the data\cite{benjamini1988}. As shown in
Figure~\ref{fig:histplot} the histplot seperates the box
into  5 bins and scales the width of that section of the box based on the
number of values in that bin; it's essentially a rotated histogram. The
vaseplot, shown in Figure~\ref{fig:vase} removes the bins and is instead based
on the estimated density at that point in the data. The specific shape is
dependent on what sort of density estimation method is used, such as kernal
estimation or nearest neighbor estimation \cite{chambers1983}. The
violinplot \cite{hintz1998} expands this concept further by representing the
width using the density-trace method proposed by Chambers
et. al \cite{chambers1983}, as seen in Figure~\ref{fig:violinplot}%%is density
%%trace the same as a pdf?
The letter-value plot\cite{hofmann} ....

@@

\begin{figure}
%%\includegraphics{bagplot}
\end{figure}
The boxplot is inherently limited to 1D data; the bagplot was introduced to
visualize 2D and higher data \cite{roussow99}.

\begin{figure}
%%\includegraphics
\end{figure}
While boxplots were designed to visualize discrete observations, often there
are no discrete observations and instead the `observation` is a timeseries
curve (where the connections are as important as the individual points) or
spatial region. To visualize these datasets, Hyndman and Shang proposed the
rainbow plot, functional bagplot and high density region (HDR)
plot\cite{Hyndman2009}. These visualizations rely on the idea of ordering the
functions such that there's some measure of how close these functions are to
each other-functional depth\cite{fabrero2007}, bivariate score depth, and
bivariate kernel density estimation\cite{scott92} are some examples. This
ordering then yields functional analogs to median and other percentile
metrics. The rainbow plot simply colors the data based on its depth, shown
in \ref{fig:rainbow}

\begin{figure}
%%\includegraphics{abc}
\end{figure}

@@

\begin{figure}
\includegraphics{figs/countour_weather.png}
\label{fig:countour}
\end{figure}
Countour Boxplots \cite{whitaker13} are an extension of functional boxplots that tries to better encapsulate outliers. Box plots inherently clip the uncertainity they show to some upper and lower band, 
creating a somewhat bounding envelope for the function that looks fairly
regular when aggregated. Countour boxplots aim instead to capture the
variablity of the uncertainity via a band depth ordering metric. Each ensemble members band depth is computed as sum of the probabilities that the observations in any given ensemble fall within the max-min envelope defined by any two other ensembles. The bands are then sorted by band depth such that the median is the ensemble with about 50\% of its members withen all envelopes formed by other bands (so most centered). Outer bands are chosen according to the task at hand. In \cite{whitaker13}, they apply the countour boxplots to temperature visualizations. %
The authors argue that countour boxplots are an improved visual idiom over the traditional spagehttie plots seen in \ref{fig:countour} because as seen in (fig), the spagehetti plots get nosy as the number of plots increase and so it's hard to tease out specific patterns. Instead in (fig3), the authors remove most of the bands and instead visualize an outliser envelope(light gray) and a more central envelope(dark gray). It retains much of the information of the spagethhit, but removes the visual noise of the lines.

>>>>

Because the box-plot does not give any indication of how data is distributed
inside the box or along the whiskers, the histplot and vaseplot were introduced to indicate
the density of the data\cite{benjamini1988}. As shown in
Figure~\ref{fig:histplot} the histplot seperates the box
into  5 bins and scales the width of that section of the box based on the
number of values in that bin; it's essentially a rotated histogram. The
vaseplot, shown in Figure~\ref{fig:vase} removes the bins and is instead based
on the estimated density at that point in the data. The specific shape is
dependent on what sort of density estimation method is used, such as kernal
estimation or nearest neighbor estimation \cite{chambers1983}. The
violinplot \cite{hintz1998} expands this concept further by representing the
width using the density-trace method proposed by Chambers
et. al \cite{chambers1983}, as seen in Figure~\ref{fig:violinplot}%%is density
%%trace the same as a pdf?
%%The letter-value plot\cite{hofmann} ....

@@

\begin{figure}
%%\includegraphics{bagplot}
\end{figure}
The boxplot is inherently limited to 1D data; the bagplot was introduced to
visualize 2D and higher data \cite{roussow1999}.
Bivariate HDR plot \cite{Hyndman1996}

While boxplots were designed to visualize discrete observations, often there
are no discrete observations and instead the `observation` is a timeseries
curve (where the connections are as important as the individual points) or
spatial region. These visualizations rely on the idea of ordering the
functions such that there's some measure of how close these functions are to
each other-functional depth\cite{fabrero2007}, bivariate score depth\cite{Hyndman2009}, and
bivariate kernel density estimation\cite{scott92} are some examples. This
ordering then yields functional analogs to median and other percentile
metrics. For small values, the ordered curves can then be colored based on the
ordering metric \cite{Hyndman2009}, but this method is unweildy for larger
ensembles.

\begin{figure}
%%\includegraphics{funcbaghdr}
\label{fig:funcbag}
\end{figure}
Instead, Hyndmand and Shang proposed functional versions of the bag and HDR
plots\cite{Hyndman2009}, as shown in Figure~\ref{fig:funcbag}. In these graphs,
what was traditionally the box part of a box plot is now a region bounded by
the curves (obtained by using one of the previously mentioned ordering
metrics). The median curve is defined as a blackline and the exceptional
outlines are shown. By using regions defined by the curves, the functional bag
and HDR plots display the variance in the ensembles while simulatanously
flattening the similaraties so that the outliers are strongly highlighted. The
HDR method better highlights outliers closer to the median, whereas the bagplot
method better highlights the outliers outside the envelopes of functions.

@@

\begin{figure}
\includegraphics{figs/countour_weather.png}
\label{fig:countour}
\end{figure}
While functional bag, box, and HDR plots all show outliers, each method is most
successful visualizing a specific type of outlier. Countour
boxplots\cite{whitaker3} aim instead to capture the variablity of the uncertainity via a band depth ordering metric. Each ensemble members band depth is computed as sum of the probabilities that the observations in any given ensemble fall within the max-min envelope defined by any two other ensembles. The bands are then sorted by band depth such that the median is the ensemble with about 50\% of its members withen all envelopes formed by other bands (so most centered). Outer bands are chosen according to the task at hand. In \cite{whitaker13}, they apply the countour boxplots to temperature visualizations. %
The authors argue that countour boxplots are an improved visual idiom over the traditional spagehttie plots seen in \ref{fig:countour} because as seen in (fig), the spagehetti plots get nosy as the number of plots increase and so it's hard to tease out specific patterns. Instead in (fig3), the authors remove most of the bands and instead visualize an outliser envelope(light gray) and a more central envelope(dark gray). It retains much of the information of the spagethhit, but removes the visual noise of the lines.
